---
title: "MUSA500 - HW1"
author: "Chenxi Zhu, Teresa Chang, and Tiffany Luo "
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 80
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

## 2. Methods

### 2.1 Data Cleaning

### 2.2 Exploratory Data Analysis

### 2.3 Multiple Regression Analysis

### 2.4 Additional Analyses

### 2.5 Software

## 3. Results

```{r load_packages, warning = FALSE, include=FALSE}
options(scipen=10000000)

library(tidyverse)
library(kableExtra)
library(gridExtra)
library(caret)
library(knitr) 
library(pscl)
library(plotROC)
library(pROC)
library(lubridate)
library(scales)
library(stargazer)
library(cowplot)
library(ggpubr)
library(sf)
```

### 3.1 Exploratory Results

#### 3.1.1 Summary Statistics

```{r}
data <- read.csv("RegressionData.csv")
```

The summary statistics reveals a diverse and complex socio-economic and housing
landscape. The significant variability in median house values suggests a wide
range of property prices, reflecting diverse regional housing markets and
economic conditions. High standard deviations in the number of households living
in poverty and the percentage of individuals with bachelor's degrees or higher
indicate notable socio-economic disparities across different census block group.

```{r}
MedHouseValue = c('Median House Value', mean(data$MEDHVAL), sd(data$MEDHVAL))
HH_poverty = c('# Households Living in Poverty', mean(data$NBELPOV100), sd(data$NBELPOV100))
bachelor = c('% with Bachelor’s Degrees or Higher', mean(data$PCTBACHMOR), sd(data$PCTBACHMOR))
vacant = c('% of Vacant Houses', mean(data$PCTVACANT), sd(data$PCTVACANT))
SingleHouse = c('% of Single House Units', mean(data$PCTSINGLES), sd(data$PCTSINGLES))

table = as.data.frame(t(data.frame(
              MedHouseValue,
              HH_poverty,
              bachelor,
              vacant,
              SingleHouse))) 

# TODO: rounding the digits
colnames(table) = c("Variable", "Mean", "SD")
table$Mean = as.numeric(table$Mean)
table$SD = as.numeric(table$SD)
table = table %>% mutate(across(where(is.numeric), round, digits=2))

table %>%
  kbl(caption = "Summary Statistics") %>%
  kable_styling() %>%
  kable_classic(html_font = "Cambria", position = "left", full_width = F)

```

#### 3.1.2 Distribution

As observed from the histograms, all variables are not normally distributed by
right skewed. By transforming the variables using `log([VAR])` (or
`log(1+[VAR])` if the variable has any zero values), some distributions of the
transformed variables look normal while some have a large spike at zero. Based
on the comparison, the subsequent analysis and the regression model will use

-   the log-transformed variables `LNMEDHVAL` and `LNNBELPOV100`

-   and the original variables `PCBACHMORE`, `PCTVACANT`, and `PCTSINGLES`.

```{r fig.height = 9, fig.width = 9}
original_vars <- c("MEDHVAL", "PCTBACHMOR", "NBELPOV100", "PCTVACANT", "PCTSINGLES")

for (var in original_vars) {
  log_var_name <- paste0("LN", var)
  data[[log_var_name]] <- ifelse(data[[var]] == 0, log1p(data[[var]]), log(data[[var]]))
}

# Function to plot histogram
plot_histogram <- function(data, var, title) {
  ggplot(data, aes_string(x = var)) + 
    geom_histogram(bins = 25, fill = "skyblue", color = "dark gray") +
    theme_minimal() +
    ggtitle(title) +
    theme(plot.title = element_text(hjust = 0.5))
}

# Plotting histograms
plot_list <- list()
for (var in original_vars) {
  plot_title <- table$Variable[which(original_vars  == var)]
  original_plot <- plot_histogram(data, var, sprintf("Histogram of %s", plot_title))
  log_plot <- plot_histogram(data, paste0("LN", var), sprintf("Histogram of LN %s", plot_title))
  plot_list[[length(plot_list) + 1]] <- original_plot
  plot_list[[length(plot_list) + 1]] <- log_plot
}

# Arrange the plots in a grid
do.call("grid.arrange", c(plot_list, ncol=2))

```

In addition to normality, the other regression assumptions will be examined in a
separate section below (Regression Assumption Checks).

b.  Look at whether the relationship between the dependent variable and each of
    the predictors is linear. present all four scatter plots as a single figure.

```{r}
# create a new plotting window and set the plotting area into a 2*2 array
par(mfrow = c(2, 2))

plot(data$LNMEDHVAL, data$LNNBELPOV100)
plot(data$LNMEDHVAL, data$PCTBACHMOR)
plot(data$LNMEDHVAL, data$PCTVACANT)
plot(data$LNMEDHVAL, data$PCTSINGLES)
```

```{r}

cor(data$LNMEDHVAL, data$LNNBELPOV100, method = c("pearson"), use = "complete.obs")
cor(data$LNMEDHVAL, data$PCTBACHMOR, method = c("pearson"), use = "complete.obs")
cor(data$LNMEDHVAL, data$PCTVACANT, method = c("pearson"), use = "complete.obs")
cor(data$LNMEDHVAL, data$PCTSINGLES, method = c("pearson"), use = "complete.obs")

```

```{r}

A <- ggscatter(data, x = "LNMEDHVAL", y = "LNNBELPOV100", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "LNMEDHVAL", ylab = "LNNBELPOV100")

B <- ggscatter(data, x = "LNMEDHVAL", y = "PCTBACHMOR", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "LNMEDHVAL", ylab = "PCTBACHMOR")


C <- ggscatter(data, x = "LNMEDHVAL", y = "PCTVACANT", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "LNMEDHVAL", ylab = "PCTVACANT")

D <- ggscatter(data, x = "LNMEDHVAL", y = "PCTSINGLES", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "LNMEDHVAL", ylab = "PCTSINGLES")

ggarrange(A, B, C, D + rremove("x.text"),  heights = c(10, 10),
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2)
```

#### 3.1.3 Choropleth Maps

iii. Present the choropleth maps of the dependent variable and the predictors.
iv. Refer to the maps in the text, and talk about the following:
v.  Which maps look similar? Which maps look different? That is, which
    predictors do you expect to be strongly associated with the dependent
    variable based on the visualization? Also, given your examination of the
    maps, are there any predictors that you think will be strongly
    inter-correlated? That is, do you expect severe multicollinearity to be an
    issue here? Discuss this in a paragraph.

#### 3.1.4 Correlation Matrix

iv. Present the correlation matrix of the predictors which you obtained from R.
v.  Talk about whether the correlation matrix shows that there is severe
    multicollinearity.
vi. Does the correlation matrix support your conclusions based on your visual
    comparison of predictor maps?

```{r}
#data %>% 
 # filter_if(~is.numeric(.), all_vars(!is.infinite(.)))
data <- data %>% 
  filter_all(all_vars(!is.infinite(.)))

cormatrix <- data[, c(4,6,7,11)]
res <- cor(cormatrix)
round(res, 2)
```

```{r}

library(corrplot)
corrplot(res, type = "upper", 
         tl.col = "black", tl.srt = 45)
```

### 3.2 Regression Results

i.  Present the regression output from R. Be sure that your output presents the
    parameter estimates (and associated standard errors, t-statistics and
    p-values), as well as the R2, the adjusted R2, and the relevant F-ratio and
    associated p-value.

ii. Referencing the regression output in (i) above, interpret the results as in
    the example included above this report outline.

NOTE: YOUR DEPENDENT VARIABLE (AND SOME PREDICTORS) WOULD BE LOG-TRANSFORMED,
UNLIKE IN THE EXAMPLE HERE. LOOK AT THE SLIDES FOR EXAMPLES OF INTERPRETING
REGRESSION OUTPUT WITH LOG-TRANSFORMED VARIABLES.

### 3.3 Regression Assumption Checks

i.  First state that in this section, you will be talking about testing model
    assumptions. State that you have already looked at the variable
    distributions earlier.

ii. Present scatter plots of the dependent variable and each of the predictors.
    State whether each of the relationships seems to be linear, as assumed by
    the regression model. [Hint: they will not look linear.]

iii. Present the histogram of the standardized residuals. State whether the
     residuals look normal.

iv. Present the 'Standardized Residual by Predicted Value' scatter plot. What
    conclusions can you draw from that? Does there seem to be
    heteroscedasticity? Do there seem to be outliers? Anything else? Discuss.

<!-- -->

1.  Mention what standardized residuals are.

<!-- -->

v.  Referencing the maps of the dependent variable and the predictors that you
    presented earlier, state whether there seems to be spatial autocorrelation
    in your variables. That is, does it seem that the observations (i.e., block
    groups) are independent of each other? Briefly discuss.

<!-- -->

vi. Now, present the choropleth map of the standardized regression residuals. Do
    there seem to be any noticeable spatial patterns in them? That is, do they
    seem to be spatially autocorrelated?

<!-- -->

1.  You will examine the spatial autocorrelation of the variables and residuals
    and run spatial regressions in the next assignment.

### 3.4 Additional Models

i.  Present the results of the stepwise regression and state whether all 4
    predictors in the original model are kept in the final model.

ii. Present the cross-validation results -- that is, compare the RMSE of the
    original model that includes all 4 predictors with the RMSE of the model
    that only includes PCTVACANT and MEDHHINC as predictors.

## 4. Discussion and Limitations
